### 2014-01-09 manuscripts/dpp-msbayes/dpp-msbayes-main.tex
Adding divergence model plots to dpp ms.

\subsection*{An empirical application}
We also assessed the behavior of the newly implemented models when applied to
the empirical dataset of \citet{Oaks2012} comprised of sequence data from 22
pairs of taxa from the Philippine Islands (Drayad DOI: 10.5061/dryad.5s07m).
We analyzed these data under five different models, which are detailed in
Table~\ref{tabEmpiricalModels}.
All of these models except one (\empModelDPPSimple) have six free demographic
parameters per pair of taxa (
\ancestralTheta{},
\descendantTheta{1}{},
\descendantTheta{2}{},
\bottleTime{},
\bottleScalar{1}{}, and
\bottleScalar{2}{}),
in addition to the $\popSampleSize{i}{} - 1$ coalescent times.
Three of these models use a Dirichlet-process prior on divergence models:
\empModelDPP, \empModelDPPInform, and \empModelDPPSimple.
The \empModelDPP model is meant to represent the prior that \citet{Oaks2012}
would have selected to represent their prior uncertainty about the parameters
of the model if provided the more flexible distributions that are now
implemented.
To assess prior sensitivity, the \empModelDPPInform model uses a more
informative exponentially distributed prior on divergence times, but otherwise
is identical to \empModelDPP.
To assess sensitivity to parameterization, we also applied the simplest
possible model under the new implementation (\empModelDPPSimple) with only a
single demographic parameter (\myTheta{}) per taxon pair, in addition to the
$\popSampleSize{i}{} - 1$ coalescent times.
We also applied the original \msBayes model (\empModelOld) with priors
selected to make it comparable to the \empModelDPP model;
the uniform prior on divergence times was selected to have the same variance as
the exponential prior of the \empModelDPP, and the mean prior on population
size was selected to have the same mean so that the models are on the same
timescale.
We also applied a model with a uniform distribution over divergence models
(\empModelUniform).
For each of these models, we simulated 2\e{7} samples from the prior, and
retained an approximate posterior of the 10,000 samples with the smallest
Euclidean distance from the summary statistics calculated from the empirical
sequence alignments.

>>>>

\subsection*{An empirical application}
We also assessed the behavior of the newly implemented models when applied to
the empirical dataset of \citet{Oaks2012} comprised of sequence data from 22
pairs of taxa from the Philippine Islands (Drayad DOI: 10.5061/dryad.5s07m).
We analyzed these data under five different models, which are detailed in
Table~\ref{tabEmpiricalModels}.
All of these models except one (\empModelDPPSimple) have six free demographic
parameters per pair of taxa (
\ancestralTheta{},
\descendantTheta{1}{},
\descendantTheta{2}{},
\bottleTime{},
\bottleScalar{1}{}, and
\bottleScalar{2}{}),
in addition to the $\popSampleSize{i}{} - 1$ coalescent times.
Three of these models use a Dirichlet-process prior on divergence models:
\empModelDPP, \empModelDPPInform, and \empModelDPPSimple.
The \empModelDPP model is meant to represent the prior that \citet{Oaks2012}
would have selected to represent their prior uncertainty about the parameters
of the model if provided the more flexible distributions that are now
implemented.
To assess prior sensitivity, the \empModelDPPInform model uses a more
informative exponentially distributed prior on divergence times, but otherwise
is identical to \empModelDPP.
To assess sensitivity to parameterization, we also applied the simplest
possible model under the new implementation (\empModelDPPSimple) with only a
single demographic parameter (\myTheta{}) per taxon pair, in addition to the
$\popSampleSize{i}{} - 1$ coalescent times.
We also applied the original \msb model (\empModelOld) with priors
selected to make it comparable to the \empModelDPP model;
the uniform prior on divergence times was selected to have the same variance as
the exponential prior of the \empModelDPP, and the mean prior on population
size was selected to have the same mean so that the models are on the same
timescale.
We also applied a model with a uniform distribution over divergence models
(\empModelUniform).
For each of these models, we simulated 2\e{7} samples from the prior, and
retained an approximate posterior of the 10,000 samples with the smallest
Euclidean distance from the summary statistics calculated from the empirical
sequence alignments.

### 2014-01-09 manuscripts/dpp-msbayes/dpp-msbayes-si-figures.tex
Working on DPP ms.
Finishing incorporation of the new figures. Also, updating some of the
new figures to add an extra digit for the exponential distributions of
divergence times.
Adding divergence model plots to dpp ms.

%% Power dispersion results
\siSidewaysFigure{../../power-comparison/images/old-power-omega-6.pdf}{
    \powerComment{\powerSeriesOld}
    \powerDispersionComment{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiOld}

\siSidewaysFigure{../../power-comparison/images/uniform-power-omega-6.pdf}{
    \powerComment{\powerSeriesUniform}
    \powerDispersionComment{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiUniform}

\siSidewaysFigure{../../power-comparison/images/exp-power-omega-6.pdf}{
    \powerComment{\powerSeriesExp}
    \powerDispersionComment{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiExp}

%% Power psi probability results
\siSidewaysFigure{../../power-comparison/images/old-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesOld}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiOld}

\siSidewaysFigure{../../power-comparison/images/uniform-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesUniform}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiUniform}

\siSidewaysFigure{../../power-comparison/images/exp-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesExp}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiExp}

%% Power dispersion prob results
\siSidewaysFigure{../../power-comparison/images/old-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesOld}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiOld}

\siSidewaysFigure{../../power-comparison/images/uniform-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesUniform}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiUniform}

\siSidewaysFigure{../../power-comparison/images/exp-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesExp}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiExp}

>>>>

%% Power dispersion results
\siSidewaysFigure{../../power-comparison/images/old-power-omega-6.pdf}{
    \powerComment{\powerSeriesOld}
    \powerDispersionComment{\powerSeriesOld}
    \timeConversionComment
}{figPowerOmegaOld}

\siSidewaysFigure{../../power-comparison/images/uniform-power-omega-6.pdf}{
    \powerComment{\powerSeriesUniform}
    \powerDispersionComment{\powerSeriesUniform}
    \timeConversionComment
}{figPowerOmegaUniform}

\siSidewaysFigure{../../power-comparison/images/exp-power-omega-6.pdf}{
    \powerComment{\powerSeriesExp}
    \powerDispersionComment{\powerSeriesExp}
    \timeConversionComment
}{figPowerOmegaExp}

%% Power psi probability results
\siSidewaysFigure{../../power-comparison/images/old-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesOld}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
}{figPowerPsiProbOld}

\siSidewaysFigure{../../power-comparison/images/uniform-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesUniform}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesUniform}
    \timeConversionComment
}{figPowerPsiProbUniform}

\siSidewaysFigure{../../power-comparison/images/exp-power-psi-prob-6.pdf}{
    \powerSupportComment{\powerSeriesExp}
    \powerProbComment{$p(\divTimeNum = 1 | \ssSpace)$}{\powerSeriesExp}
    \timeConversionComment
}{figPowerPsiProbExp}

%% Power dispersion prob results
\siSidewaysFigure{../../power-comparison/images/old-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesOld}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesOld}
    \timeConversionComment
}{figPowerOmegaProbOld}

\siSidewaysFigure{../../power-comparison/images/uniform-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesUniform}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesUniform}
    \timeConversionComment
}{figPowerOmegaProbUniform}

\siSidewaysFigure{../../power-comparison/images/exp-power-omega-prob-6.pdf}{
    \powerSupportComment{\powerSeriesExp}
    \powerProbComment{$p(\divTimeDispersion < 0.01 | \ssSpace)$}{\powerSeriesExp}
    \timeConversionComment
}{figPowerOmegaProbExp}

% Empirical results
\siEightFigure{../../empirical-analyses/plots/philippines-dpp-div-models.pdf}{
    The divergence-model results when the 22 pairs of taxa from the Philippines
    are analyzed under the \empModelDPP model (Table~\ref{tabEmpiricalModels}).
    The 10 unordered divergence models with highest posterior probability
    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
    indicate the inferred number of taxon pairs that diverged at each event.
    The times indicate the posterior median and 95\% highest posterior density
    (HPD) interval conditional on each divergence model.
    For each model, times are summarized across posterior samples by the number
    of taxon pairs associated with each divergence. For models in which there
    are multiple divergence events with the same number of taxon pairs, the
    events are sorted by time to summarize the divergence times in a consistant
    way.
}{figDivModelsDPP}

\siEightFigure{../../empirical-analyses/plots/philippines-dpp-inform-div-models.pdf}{
    The divergence-model results when the 22 pairs of taxa from the Philippines
    are analyzed under the \empModelDPPInform model
    (Table~\ref{tabEmpiricalModels}).
    The 10 unordered divergence models with highest posterior probability
    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
    indicate the inferred number of taxon pairs that diverged at each event.
    The times indicate the posterior median and 95\% highest posterior density
    (HPD) interval conditional on each divergence model.
    For each model, times are summarized across posterior samples by the number
    of taxon pairs associated with each divergence. For models in which there
    are multiple divergence events with the same number of taxon pairs, the
    events are sorted by time to summarize the divergence times in a consistant
    way.
}{figDivModelsDPPInform}

\siEightFigure{../../empirical-analyses/plots/philippines-dpp-simple-div-models.pdf}{
    The divergence-model results when the 22 pairs of taxa from the Philippines
    are analyzed under the \empModelDPPSimple model
    (Table~\ref{tabEmpiricalModels}).
    The 10 unordered divergence models with highest posterior probability
    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
    indicate the inferred number of taxon pairs that diverged at each event.
    The times indicate the posterior median and 95\% highest posterior density
    (HPD) interval conditional on each divergence model.
    For each model, times are summarized across posterior samples by the number
    of taxon pairs associated with each divergence. For models in which there
    are multiple divergence events with the same number of taxon pairs, the
    events are sorted by time to summarize the divergence times in a consistant
    way.
}{figDivModelsDPPSimple}

\siEightFigure{../../empirical-analyses/plots/philippines-uniform-div-models.pdf}{
    The divergence-model results when the 22 pairs of taxa from the Philippines
    are analyzed under the \empModelUniform model
    (Table~\ref{tabEmpiricalModels}).
    The 10 unordered divergence models with highest posterior probability
    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
    indicate the inferred number of taxon pairs that diverged at each event.
    The times indicate the posterior median and 95\% highest posterior density
    (HPD) interval conditional on each divergence model.
    For each model, times are summarized across posterior samples by the number
    of taxon pairs associated with each divergence. For models in which there
    are multiple divergence events with the same number of taxon pairs, the
    events are sorted by time to summarize the divergence times in a consistant
    way.
}{figDivModelsUniform}

\siEightFigure{../../empirical-analyses/plots/philippines-old-div-models.pdf}{
    The divergence-model results when the 22 pairs of taxa from the Philippines
    are analyzed under the \empModelOld model
    (Table~\ref{tabEmpiricalModels}).
    The 10 unordered divergence models with highest posterior probability
    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
    indicate the inferred number of taxon pairs that diverged at each event.
    The times indicate the posterior median and 95\% highest posterior density
    (HPD) interval conditional on each divergence model.
    For each model, times are summarized across posterior samples by the number
    of taxon pairs associated with each divergence. For models in which there
    are multiple divergence events with the same number of taxon pairs, the
    events are sorted by time to summarize the divergence times in a consistant
    way.
}{figDivModelsOld}

\siEightFigure{../../empirical-analyses/plots/negros-panay-dpp-unordered-div-models.pdf}{
    The divergence-model results when the 9 pairs of taxa from the Islands of
    Negros and Panay are analyzed under the \npModelDPP model sampling over
    unordered models of divergence.
    (Table~\ref{tabEmpiricalModels}).
    The 10 unordered divergence models with highest posterior probability
    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
    indicate the inferred number of taxon pairs that diverged at each event.
    The times indicate the posterior median and 95\% highest posterior density
    (HPD) interval conditional on each divergence model.
    For each model, times are summarized across posterior samples by the number
    of taxon pairs associated with each divergence. For models in which there
    are multiple divergence events with the same number of taxon pairs, the
    events are sorted by time to summarize the divergence times in a consistant
    way.
}{figDivModelsNP}

\siEightFigure{../../empirical-analyses/plots/negros-panay-dpp-ordered-div-models.pdf}{
    The divergence-model results when the 9 pairs of taxa from the Islands of
    Negros and Panay are analyzed under the \npModelDPPOrdered model sampling
    over ordered models of divergence.
    (Table~\ref{tabEmpiricalModels}).
    The posterior sample of divergence models were summarized while ignoring
    the identity of the taxon pairs in order to compare the results of the
    \npModelDPP model.
    The 10 unordered divergence models with highest posterior probability
    ($p(\divTimeIndexVector \given \ssSpace)$) are shown, where the numbers
    indicate the inferred number of taxon pairs that diverged at each event.
    The times indicate the posterior median and 95\% highest posterior density
    (HPD) interval conditional on each divergence model.
    For each model, times are summarized across posterior samples by the number
    of taxon pairs associated with each divergence. For models in which there
    are multiple divergence events with the same number of taxon pairs, the
    events are sorted by time to summarize the divergence times in a consistant
    way.
}{figDivModelsNPOrdered}

\siFigure{../../empirical-analyses/plots/negros-panay-marginal-times.pdf}{
    The marginal divergence-time results when the 9 pairs of taxa from the
    Islands of Negros and Panay are analyzed under the \npModelDPPOrdered model
    that samples over ordered models of divergence
    (Table~\ref{tabEmpiricalModels}).
}{figMarginalTimes}
